---
title: "R_bootcamp_day2"
author: "Tanya Tandon"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
here::here("data", "gapminder5.csv")

gapminder <- read.csv(here::here("data", "gapminder5.csv"))
str(gapminder)


```
# Converting factors to characters 
```{r}

gapminder$country <- as.character(gapminder$country)
gapminder$continent <- as.character(gapminder$continent)
str(gapminder)




```
#GDP for each country 
```{r}

obs <- 1:nrow(gapminder)

for(i in obs) {
  gapminder[i, "gdp"] <- gapminder[i, "pop"] * gapminder[i, "gdpPercap"]
}

```
# a new variable that finds that natural log (log) of the GDP per capita and of population - call them log_gdpPercap and log_pop

```{r}
for (i in nrow(gapminder)){
  log_gdpPercap <- log(gapminder$gdpPercap)
  log_pop <- log(gapminder$pop)
  head(gapminder)
}


```

```{r}

for (i in obs) { # the function to repeat is enclosed in braces {}
  gapminder[i, "log_gdpPercap"] <- log(gapminder[i, "gdpPercap"])
  gapminder[i, "log_pop"] <- log(gapminder[i, "pop"])
}
```

```{r}
gapminder$vec_log_gdpPercap <- log(gapminder$gdpPercap)
all(gapminder$vec_log_gdpPercap == gapminder$log_gdpPercap)
```
# Has life expectancy increased over time in continent

```{r}
continent1 <- unique(gapminder$continent)
for ( i in continent1) {
  mean_conti <- mean(gapminder$lifeEx[gapminder$continent == i], na.rm = T)
  print(paste0(i, ": " , mean_conti))
}
```

#nested for loop exercise!Has the gap in life expectancy between countries on different continents narrowed over time?

```{r}
un <- unique(gapminder$continent)
for( i in un ){
  gap <- sd(gapminder$country[gapminder$continent == i])
  print( paste0(i , " : " , gap))
}

```
# WHILLEE
```{r}
i <- 1987 

while(i <= 2002) {
  sd <- sd(gapminder$lifeExp[gapminder$year==i])
  print(paste( i , ": " , sd))
  i <-i + 5
}

```
#Write a for loop that reports the mean population for years greater than or equal to 1987. Make sure the loop prints a message if the condition is not met!

```{r}
gapminder
 for(i in unique(gapminder$year)){
   m  <- mean(gapminder$pop[gapminder$year == i])
   if (i >= 1987) {
     print(paste0(" Pop in ", i, " : ", m))
   }
   else {
     print( " Year smaller than 1987")
   }
   
 }
```
```{r}
report_mean_sd <- 
  function(df, variable, country) {
    var <- gapminder[[variable]][gapminder$country == country]
    head(var)
  }

```