---
title: "Day3RGGplot_TT"
author: "Tanya Tandon"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
source(here::here("data/day3_objects.R"))
```
#Using the gapminder07 data, create a scatterplot of the natural log of gdpPercap as a function of the natural log of  pop. Give it a title and axis labels.

```{r}
ggplot(gapminder07) + geom_point(aes(  y = log(pop), x= log(gdpPercap), )) + labs(x = "log gdpPercap", y = "log pop", title = "scatter plot" )
```
#Plot a column chart hydroelectric power generated over time.
```{r}
merged_energy %>%
  group_by(datetime) %>%
  summarise(total= )
  
```

#Create a column chart that shows the total output per source. Change the color of the columns to "darkred". Add a horizontal line indicating the mean output across all sources. Use the cheatsheet to identify the geom function that you need.Add a meaningful title and axis labels using labs().

```{r}
long_merged_energy %>%
  group_by(source) %>%
  summarise(total_output = sum(output)) %>%
  ggplot() + geom_col(aes (x= source, y= total_output ), col = "darkred", fill = "darkred") + geom_hline()
  
  
```

#Create a line plot that compares generation of wind, solar, and geothermal energy over time.

```{r}
long_merged_energy %>%
  filter(source == 'wind' | source == 'solar' | source ==?re 'geothermal') %>%
  ggplot() +
  geom_line(aes(x=datetime, y=output, group=source, col=source))
```

#Visualize the average output for each hour of the day, grouped by source.
```{r}
long_merged_energy %>%
  
  mutate ( hour = lubridate:: hour(datetime) ) %>%
   group_by(source, hour) %>% 
  summarize( mean = mean(output)) %>%
  ggplot() +
   geom_area(aes(x=hour, y=output, fill=factor(source))) + 
  scale_fill_brewer(palette="Set3", name="Source") + 
  labs(title="Average hourly output by source", 
       subtitle="Data collected during September 3-9",
       x="Hour of the day", y="Output (MW)") + 
  theme_bw()

```

```{r}
long_gen %>% 
  ggplot() + 
  geom_line(aes(x = datetime, y = output)) + 
  facet_wrap(~source, scales="free") + 
  labs(title="Generation over time, by energy source", subtitle="Hourly data from September 3-9, 2018", x="Hour", y="Output (MW)")
merge(regroup, by="type")
```

#Compare generation patterns for each source
